---
title: "IE6600"
subtitle: "<br/>k - Nearest Neighbor"
author: "Zhenyuan Lu"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css","footer.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    # yolo: 
    #   img: 'file:///E:/Data%20Analytics%20Engineering/IE6600_ZL/Materials/R/takeABreak.png'
    #   times: 1 
---

layout: true
  
<div class="my-header"></div>

<div class="my-footer"><span>© 2020 Zhenyuan Lu  
</span></div> 


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = T,eval=T, warning=F)
library(tidyverse)
library(class)
library(caret)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(base_color = "#76abd9",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono")
)

```

---
class: center, middle

# k - Nearest Neighbor


---
# Step1: Prerequisites
To read more about this dataset, refer to:
Mangasarian OL, Street WN, Wolberg WH. Breast cancer diagnosis and prognosis via linear programming. Operations Research. 1995; 43:570-577.
```{r eval=F}
library(tidyverse)
```

```{r eval=T}
wbcd <- readr::read_csv("data/kNN_wbcd.csv", guess_max=3000)
wbcd <- wbcd %>% select(-1, -X33) %>% as.data.frame()
head(wbcd)
```

---
# Step2: exploring and preparing the data
Many R machine learning classifiers require that the target feature is coded as a
factor, so we will need to recode the diagnosis variable. We will also take this
opportunity to give the "B" and "M" values more informative labels using the
labels parameter:
```{r eval=T}
table(wbcd$diagnosis)
```


```{r eval=T}
wbcd$diagnosis <- factor(
  wbcd$diagnosis,
  levels = c("B", "M"),
  labels = c("Benign", "Malignant")
)
```


---
Now, when we look at the prop.table() output, we notice that the values have
been labeled Benign and Malignant with 62.7 percent and 37.3 percent of the
masses, respectively:
```{r eval=T}
round(prop.table(table(wbcd$diagnosis)) * 100, digits = 1)
```


---
The remaining 30 features are all numeric, and as expected, they consist of three
different measurements of ten characteristics. For illustrative purposes, we will
only take a closer look at three of these features
```{r eval=T}
wbcd %>% select(radius_mean,area_mean,smoothness_mean) %>% summary()

```


---
# Transformation – normalizing numeric data

```{r eval=T}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r eval=T}
wbcd_n <- as.data.frame(lapply(wbcd[2:31], normalize))
summary(wbcd_n$area_mean)
```

---
# Data preparation – creating training and test datasets

```{r}
ran <- sample(1:nrow(wbcd), 0.9 * nrow(wbcd)) 
wbcd_train <- wbcd_n[ran,] 
wbcd_test <- wbcd_n[-ran,] 

wbcd_train_labels <- wbcd[ran, 1]
wbcd_test_labels <- wbcd[-ran, 1]
```


---
# Step 3 – training a model on the data

```{r eval=F}
library(class)
```

  ```r
  knn(train, test, class, k)
  ```

```{r}
wbcd_test_pred <- knn(
  train = wbcd_train,
  test = wbcd_test,
  cl = wbcd_train_labels,
  k = 21
)

```

---
# Step 4 – evaluating model performance

```{r}
caret::confusionMatrix(table(x = wbcd_test_labels, y = wbcd_test_pred))
```


---
# Step 5 – improving model performance
Transformation – z-score standardization

```{r}
wbcd_z <- as.data.frame(scale(wbcd[-1]))
ran <- sample(1:nrow(wbcd_z), 0.9 * nrow(wbcd_z)) 
wbcd_train <- wbcd_z[ran,] 
wbcd_test <- wbcd_z[-ran,] 
wbcd_train_labels <- wbcd[ran, 1]
wbcd_test_labels <- wbcd[-ran, 1]
```

---
```{r}
wbcd_test_pred <- knn(
  train = wbcd_train,
  test = wbcd_test,
  cl = wbcd_train_labels,
  k = 21
)
```

---
```{r}
caret::confusionMatrix(table(x = wbcd_test_labels, y = wbcd_test_pred))
```



